---
title: "Week6_Data_Manipulation"
author: "Josh Marciniak"
date: "February 13, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### 8.1 E1
## Importing the tidyverse and converting mtcars into a tibble
```{r}
library(tidyverse)
mtcars %>% data.frame 
```
# Rewriting mtcars[1:5, ] using dplyr and pipes returns all the columns of the first 5 rows of mtcars
```{r}
mtcars %>%
  as_tibble %>%
  slice(1:5)
```
#Displaying last 5 rows of mtcars
```{r}
mtcars%>%
  as_tibble%>%
  tail(5)
```
#Rewriting nrow(mtcars) using dplyr and pipes returns the number of rows in mtcars
```{r}
mtcars %>% 
  as_tibble%>%
  summarise(count=n())
```
#Selecting only the column named mpg
```{r}
mtcars %>% 
  as_tibble%>%
  select(mpg)
```
#Selecting rows where cyl = 6 from column mpg
```{r}
mtcars %>% 
  as_tibble%>%
  filter(cyl==6)%>%
  select(mpg)
```
#Selecting rows where cyl = 6 and all columns
```{r}
mtcars %>% 
  as_tibble%>%
  filter(cyl==6)
```
#Selecting rows where mpg is greater than 25 from columns mpg and cyl
```{r}
mtcars %>% 
  as_tibble%>%
  filter(mpg > 25) %>%
  select(c(mpg,cyl))
```
###8.2 E2
```{r}
diamonds %>% data.frame
```
#Displaying the first 5 rows of diamonds
```{r}
diamonds %>%
  as_tibble %>%
  slice(1:5)
```
#Counting the number of rows in diamonds
```{r}
diamonds %>%
  as_tibble %>%
  summarise(count=n())
```
#Counting the number of rows in diamonds that have a cut of "very good"" and displaying all columns
```{r}
diamonds %>%
  as_tibble %>%
  filter(cut=="Very Good")%>%
  summarise(count=n())
```
#Counting the number of rows in diamonds with a carat greater than 3.0
```{r}
diamonds %>%
  as_tibble %>%
  filter(carat>3.0)%>%
  summarise(count=n())
```
#Displaying the rows where color = D and the columns color and cut
```{r}
diamonds %>%
  as_tibble %>%
  filter(color == "D")%>%
  select(c(color, cut))
```
#Displaying the mean of the price column of diamonds
```{r}
diamonds %>%
  as_tibble %>%
  summarise(price.mean=mean(price))
```

###8.3 E3
##Question 1
```{r}
mtcars%>%
  as_tibble%>%
  group_by(cyl) %>%
  summarise(count=n())
```
##Question 2
```{r}
mtcars%>%
  as_tibble%>%
  group_by(cyl) %>%
  summarise(mpg.mean=mean(mpg),mpg.disp=mean(disp))
```
##Question 3
```{r}
diamonds%>%
  as_tibble%>%
  group_by(cut)%>%
  summarise(price.max=max(price),price.mean=mean(price),price.min=min(price))
```
##Question 4
```{r}
diamonds%>%
  as_tibble%>%
  group_by(color)%>%
  summarise(price.max=max(price),price.mean=mean(price),price.min=min(price))
```